Undirected Graph Exploration with $\Theta(\log\log n)$ Pebbles

### 2016-05-11 paper/main.tex
began rewriting text shizzle

\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{ex}{Exercise}

@@

\title{A \emph{Grimm} idea}
\subtitle{Exploring graphs with pebbles}
\author{Christoph Welzel}
\maketitle

\begin{abstract}
  In this paper we re-present and explain an approach on exploring large
  undirected graphs with the help of markers, so called pebbles, for vertices.
  We focus on the central ideas of this approach and explain why it is possible
  to traverse any graph with $n$ vertices with $\Theta(\log \log n)$ pebbles
  and the same amount of memory.
\end{abstract}

@@

\setquotestyle{english}
\section{Introduction}
Traversing large graphs is an interesting challenge. In times of the internet
web crawlers have to traverse immense graphs but also regarding complexity
theory this is an interesting problem. First we will informally introduce two
complexity classes to illustrate the importance of graph traversal from a
complexity point of view: $L$ as the class of problems that can be solved by
a deterministic Turing machine with additional memory of logarithmic amount
of the input. And $SL$ as the class of problems that can be solved with the
same space requirements but by a non-deterministic symmetrical Turing machine,
where symmetrical means that for every computation step of the Turing machine
there is a computation step to undo the previous one. As it turns out checking
if two vertices $s$ and $t$ are connected in an undirected graph is a problem
that is complete for $SL$. This problem is often refered to as
$\mathrm{USTCON}$ and \citeauthor{logspacealg} showed in \cite{logspacealg}
that $\mathrm{USTCON}\in L$ which collapsed $SL$ into $L$, hence $SL = L$.
In the following we will present the results from \cite{pebbles} where another
mechanism of graph exploration is used: with the use of pebbles which are small
pieces of memory that can be dropped on and collected from the vertex as an
agent traverses the graph. As we will show it is possible to traverse a graph
with $n$ vertices with $\mathcal{O}(\log\log n)$ memory and the same amount of
pebbles. Furthermore we will explain why there is no smaller amount of memory
or pebbles to generally obtain the same guarantee to traverse all $n$ vertices.

\section{Preliminaries}
For the course of this paper we assume the following: Firstly, the vertices can
not be distinguished by the traversing agent. But as the agent moves on a
vertex it can compute its degree. Secondly, the edges attached to the vertex
can be locally distinguished and are enumerated with the indices
$\left\{0,\dots,d_{v}-1\right\}$ where $d_{v}$ denotes the degree of the vertex
$v$. We refer to the \enquote{index} of the edge as its \emph{port}. Note that
every edge thus has two different ports and furthermore that there is no
relation between the ports is assumed. Additionally the degree of the
considered graphs are assumed to be limited by a constant $\Delta$ and that
there are only a finite amount of vertices.

\section{Pebble machines}
We are now ready to introduce the model of pebble machines. It closely
follows the notion of Turing machines but adds mechanisms to model the
movements of the agent. The presented model of pebble machines thus have
\begin{enumerate}
  \item a finite set of states $Q$
  \item a finite set of final states $F\subseteq Q$
  \item a finite set of distinguishable pebbles $P$
  \item a length of the used memory tape $m$
  \item a computation function $\delta$
  \item a function that describes the movement onto a vertex
    $\din$
  \item a function that describes the leaving of a vertex $\dout$
  \item an inital state $q_{0}$
\end{enumerate}
where formally the computation function models the behaviour of the pebble
machine as follows:
\begin{equation*}
  \delta\colon (Q\setminus F)\times\left\{0,1\right\}\rightarrow
  Q\times\left\{0,1\right\}\times\left\{L,R\right\}
\end{equation*}
with the typical understanding of the computation of a Turing machine.
As an agent traverses a vertex it is aware of the pebbles it carries
around, it observes the set of pebbles on the vertex and the degree of the
vertex and it furthermore aware of the port that it entered through.
Thus we use
\begin{equation*}
  \din\colon Q\times 2^{P}\times 2^{P}\times
  \left\{0,\dots,\Delta - 1\right\}\times\left\{0,\dots,\Delta - 1\right\}
  \rightarrow Q
\end{equation*}
to model this behaviour. And finally as the agent leaves a vertex it leaves
a set of pebbles behind, picks up another set and chooses the port it leaves
through. We choose to model this behaviour as
\begin{equation*}
  \dout\colon Q
  \rightarrow 2^{P}\times 2^{P}\times \left\{0,\dots,\Delta - 1\right\}
\end{equation*}
Of course we also demand consistency of the pebbles, i.e. that if the agent
carried $P_{1}\subseteq P$ and found $P_{2}\subseteq P$ it is necessary that
$P_{1}$ and $P_{2}$ are disjoint and furthermore that if the agent leaves
carrying $P'_{1}$ and leaving $P'_{2}$ behind those sets are disjoint and
$P_{1}\cup P_{2} = P'_{1}\cup P'_{2}$.

\section{Graph exploration}
In order to describe the movement of an agent we introduce the notion of an
exploration sequence as $e_{0}, e_{1},\dots$ which models relative movements
as follows. We use a sequence of variables $l_{0},l_{1},\dots$ which denotes
the port a vertex was entered through. And the next vertex is determined as the
one connected to the port $(l_{i} + e_{i}) \mod d_{v_{i}}$. We define
$l_{0} = 0$ and the notion $(e_{0},\dots,e_{k})^{\ast}$ for periodic sequences.
It can be shown that a periodic sequence produces a closed walk through a
graph.
\begin{proof}
  We argue that there is only a finite amount of triples of the
  form $(v,l,i)$ where currently the agent is on vertex $v$, entered
  this vertex through port $l$ and is currently at index $i$ of the exploration
  sequence. For two triples $(v',l',i')$ and $(v'',l'',i'')$ that both lead to
  $(v,l,i)$ it is easy to see, that $(v',l',i') = (v'',l'',i'')$. Thus, for
  every triple there is a unique predecessor and successor which implies due to
  the finiteness of the triples $(v,l,i)$ and because the agent moves
  infinitely that the first triple has to be valid infinitly often.
  This means a periodic exploration sequence yields a closed walk.
\end{proof}

From the following result of \cite{logspacealg} we present how we can obtain
an $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine that finds a closed
walk of at least $n$ distinct vertices in any graph with bounded degree.
\begin{thm}
  \label{thm:logalg}
There is an algorithm $M\in L$ which produces a universal exploration sequence
for every regular graph with $n$ vertices.
\end{thm}
First we argue that we can use this algorithm to find a universal exploration
sequence of the form $e=(e_{0},\dots,e_{k})^{\ast}$ that traverses $n$ distinct
vertices for any 3-regular graph and requires only $\mathcal{O}(\log n)$
memory. Therefore we rely on the traversal sequence of $M$ from Theorem
\ref{thm:logalg} and take from that the first $k:=3n\cdot c$ indices where
$c$ denotes the number of configurations of $M$. We can now be assured that
$e$ actually is a universal traversal sequence for any 3-regular graph because
otherwise $M$ would have already looped in a strict subset of vertices of some
3-regular graph. This contradicts that $M$ produces a universal exploration
sequence for any graph with $n$ vertices.

We can now find the $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine we
are looking for.
\begin{thm}
  \label{thm:pebblewalk}
  There is a $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine which
  explores every given graph $G$ with bounded degree $\Delta$ either completly
  or at least loops on $n$ distinct vertices.
\end{thm}
\begin{wrapfigure}{r}{0.5\textwidth}
  \caption{Illustration how to transform any graph to a 3-regular graph.}
  \label{fig:3reg}
  \resizebox{0.5\textwidth}{!}{\input{tikz/regsim}}
\end{wrapfigure}
\begin{proof}
  The main idea is to use the adapted algorithm $A$ to find for an
  universal exploration sequence for a 3-regular graph from above. Therefore
  we take $G$ and transform it to a 3-regular graph on which $A$ finds the
  corresponding universal exploration sequence. The transformation replaces
  any node $v$ with degree $d$ by $3d$ nodes which then form a circle where
  every nodes has degree $3$. The edges of $v$ are moved to nodes in the
  circle. Specifically the edge with port $i$ is replaced by 3 edges which are
  attached to $(v,i), (v,i+d), (v,i+2d)$. (Note that the other
  vertex attached to the edge is replaced by the same construction and therfore
  offers enough ports for the newly introduced edges.) Figure \ref{fig:3reg}
  illustrates this construction.  This simulation yields at most
  $3\Delta n^{2}$ many nodes and can be emulated with $\mathcal{O}(1)$ many
  memory space since it is only needed to store the second component of the
  inserted nodes which is bounded by $3\Delta$. Furthermore $A$ can be run on
  this constructed Graph and only needs $\mathcal{0}(\log n)$ space since
  $\mathcal{O}(\log(3\Delta n^{2})) = \mathcal{O}(\log(3) + \log(\Delta) +
  2\log(n)) = \mathcal{O}(\log n)$. Which yields the desired
  $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine.
\end{proof}

@@

\section{Barriers}
In the following we will discuss the limits of the exploration approach with
pebbles. Specifically, we will argue that we can construct 3-regular graphs
that any agent with $p$ pebbles cannot traverse. Such graphs are called
$p$-barriers.
Note that connecting a 3-regular graph might violate its 3-regularity.
Therefore we rely on having a distinguishable edge $e$ in the 3-regular graph
which we can delete to connect the vertices $e$ connected to the
\enquote{outside}. This preserves the 3-regularity. The interested reader might
then enjoy the following exercise:
\begin{ex}
  Given a $p$-barrier with $m$ vertices for a agent $A$ with $p$ pebbles.
  Construct a trap for $A$ (with $2m+4$ vertices). 
\end{ex}

>>>>

\newtheorem{thm}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{prop}{Proposition}

@@

\title{Exploring graphs with pebbles}
\subtitle{A \emph{Grimm} idea}
\author{Christoph Welzel}
\maketitle

@@

\setquotestyle{english}
\section{Introduction}

\section{Preliminaries}
For the further course we assume the following: Firstly, the considered
graphs are finite but huge. The considered agents have limited memory and can
therefore not compute the identifier of vertices. This makes all vertices
indistinguishable for the agents. Also, the degree of the vertices is bounded
by a constant $\Delta$. For every vertex $v$ we assume that the connected edges
can be locally enumerated with labels $0,\dots,d_{v} - 1$ where $d_{v}$ is the
degree of $v$. These edge-labels are called ports and every edge has one port
for both vertices it connects. Note that there is no relation between these
ports is presumed.

\section{Graph exploration}
We begin this section by formally defining our agents in the form of
\emph{pebble machines}. This definition closely follows the notion of Turing
machines but adds mechanisms to model the movements of the agent.
\begin{definition}[Pebble machine]
  A $(s,p,m)$-\emph{pebble machine} $\mathcal{M}$ is defined as a tuple
  $(Q,F,P,m,\delta,\din,\dout,q_{0})$. Here we have
  \begin{enumerate}
    \item a finite set of states $Q$ with cadinality $s$
    \item a finite set of final states $F\subseteq Q$
    \item a finite set of distinguishable pebbles $P$ with cardinality $p$
    \item the length of the available memory tape $m$
    \item a computation function $\delta$
    \item a function that describes the movement onto a vertex
      $\din$
    \item a function that describes the leaving of a vertex $\dout$
    \item an inital state $q_{0}$
  \end{enumerate}
  where formally the computation function models the behaviour of the pebble
  machine as
  \begin{equation*}
    \delta\colon (Q\setminus F)\times\left\{0,1\right\}\rightarrow
    Q\times\left\{0,1\right\}\times\left\{L,R\right\}
  \end{equation*}
  with the typical understanding of the computation of a Turing machine.
  Furthermore
  \begin{equation*}
    \din\colon Q\times 2^{P}\times 2^{P}\times
    \left\{0,\dots,\Delta - 1\right\}\times\left\{0,\dots,\Delta - 1\right\}
    \rightarrow Q
  \end{equation*}
  models how a pebble machine moves upon a vertex $v$ with
  $\din(q,P_{1},P_{2},d,\ell) = q'$ where $q$ is the current state of
  $\mathcal{M}$, $P_{1}$ the set of pebbles it carries, $P_{2}$ the set
  of pebbles located on $v$, $d$ is the degree of $v$ and $\ell$ the port
  through which $\mathcal{M}$ enters $v$. Then $\mathcal{M}$ starts its
  computation according to $\delta$ in state $q'$.

  $\dout$ is defined as
  \begin{equation*}
    \dout\colon Q
    \rightarrow 2^{P}\times 2^{P}\times \left\{0,\dots,\Delta - 1\right\}
  \end{equation*}
  where $\dout(q) = (P_{1}, P_{2}, \ell)$ defines that $\mathcal{M}$ leaves
  a vertex $v$ through port $\ell$ dropping the pebbles in $P_{1}$ onto $v$
  and carrying the pebble in $P_{2}$.

  We additionally demand consistency of pebbles, i.e. that if the agent
  carried the pebbles in $P_{1}\subseteq P$ and found the pebbles in
  $P_{2}\subseteq P$ it is necessary that $P_{1}$ and $P_{2}$ are disjoint and
  furthermore that if the agent leaves carrying $P'_{1}$ and leaving $P'_{2}$
  behind those sets are disjoint and $P_{1}\cup P_{2} = P'_{1}\cup P'_{2}$.
\end{definition}

In order to describe the movement of an agent we introduce the notion of an
exploration sequence as $e_{0}, e_{1},\dots$ which define relative movements:
Let $\ell = \ell_{0},\ell_{1},\dots$ be a sequence of variables where
$\ell_{i}$ denotes the port which a vertex $v$ was entered through. Then an
agent following $\ell$ leaves through port
$(\ell_{i} + e_{i}) \mod d_{v}$. We define $\ell_{0} = 0$ and the notion
$(e_{0},\dots,e_{k})^{\ast}$ for periodic exploration sequence.

Beginning from the following result of \cite{logspacealg} we present how we can
obtain an $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine that walks
along a closed walk of at least $n$ distinct vertices in any graph with bounded
degree.
\begin{thm}[Reingold]
  \label{thm:logalg}
  There is an $\mathcal{O}(\log n)$-space algorithm producing a universal
  exploration sequence for any regular graph on $n$ vertices.
\end{thm}

We can now find the $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine we
are looking for.
\begin{thm}
  \label{thm:pebblewalk}
  There is a $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine which
  explores every given graph $G$ with bounded degree $\Delta$ either completly
  or at least loops on $n$ distinct vertices.
\end{thm}
\begin{wrapfigure}{r}{0.5\textwidth}
  \caption{Illustration how to transform any graph to a 3-regular graph.}
  \label{fig:3reg}
  \resizebox{0.5\textwidth}{!}{\input{tikz/regtrans}}
\end{wrapfigure}
\begin{proof}
  The main idea is to use the adapted algorithm $A$ to find for an
  universal exploration sequence for a 3-regular graph from above. Therefore
  we take $G$ and transform it to a 3-regular graph on which $A$ finds the
  corresponding universal exploration sequence. The transformation replaces
  any node $v$ with degree $d$ by $3d$ nodes which then form a circle where
  every nodes has degree $3$. The edges of $v$ are moved to nodes in the
  circle. Specifically the edge with port $i$ is replaced by 3 edges which are
  attached to $(v,i), (v,i+d), (v,i+2d)$. (Note that the other
  vertex attached to the edge is replaced by the same construction and therfore
  offers enough ports for the newly introduced edges.) Figure \ref{fig:3reg}
  illustrates this construction.  This simulation yields at most
  $3\Delta n^{2}$ many nodes and can be emulated with $\mathcal{O}(1)$ many
  memory space since it is only needed to store the second component of the
  inserted nodes which is bounded by $3\Delta$. Furthermore $A$ can be run on
  this constructed Graph and only needs $\mathcal{0}(\log n)$ space since
  $\mathcal{O}(\log(3\Delta n^{2})) = \mathcal{O}(\log(3) + \log(\Delta) +
  2\log(n)) = \mathcal{O}(\log n)$. Which yields the desired
  $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble machine.
\end{proof}

@@

\section{Barriers}
In the following we will discuss the limits of the exploration approach with
pebbles. Specifically, we will argue that we can construct 3-regular graphs
that any agent with $p$ pebbles cannot traverse. Such graphs are called
$p$-barriers.
Note that connecting a 3-regular graph might violate its 3-regularity.
Therefore we rely on having a distinguishable edge $e$ in the 3-regular graph
which we can delete to connect the vertices $e$ connected to the
\enquote{outside}. This preserves the 3-regularity. The interested reader might
then enjoy the following exercise:
\begin{prop}
  Given a $p$-barrier with $m$ vertices for a agent $A$ with $p$ pebbles.
  Construct a trap for $A$ (with $2m+4$ vertices). 
\end{prop}

### 2016-05-11 presentation/main.tex
began rewriting text shizzle

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Exploration sequence $e_{1},\dots, e_{n}$
      \item Relative movements of agent
      \item Leaving port: $\ell_{i} + e_{i}\mod d_{v}$
      \item \emph{Universial} for a class of graphs if it explores any graph of
        that class
      \item<2->[$\rightarrow$] Example:
        $\alt<3-3>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
        \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
        \alt<5-5>{\colorbox{green}{2}}{\colorbox{white!0}{2}},
        \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \uncover<2->{
        \node[rectangle, draw=green, below right=0.3 and 1 of d1]
        {3-Regularity};
      }
      \uncover<3->{
        \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal lemma};
      }
      \uncover<4->{
        \node[rectangle, draw=yellow, below right=1.7 and 1 of d1]
        {Closed walk lemma};
      }
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{Closed walk}
  \begin{mdframed}[linecolor=yellow]
    \begin{lemma}[Closed walk]
      An agent following an exploration sequence of the form
      $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
    \end{lemma}
  \end{mdframed}
  \uncover<2->{
    \begin{mdframed}[linecolor=blue]
      \begin{lemma}[Periodic exploration sequence (PES-Lemma)]
        There exists an $\mathcal{O}(\log n)$-space algorithm producing a
        universal exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any
        $3$-regular graph with at most $n$ vertices.
      \end{lemma}
    \end{mdframed}
  }
\end{frame}

@@

\section{Pebble simulation}
\begin{frame}
  \begin{theorem}
    There is a constant $c\in\mathbb{N}$, such that for any Graph $G$
    with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$, there
    exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates the walk
    of $\mathcal{M}$ or explores $G$.
  \end{theorem}
\end{frame}

@@

\begin{frame}
  \frametitle{4. Transfer tape-content through movements}
  \begin{itemize}
    \item Simulated agent moves
    \item Preserve tape-content:
      \begin{itemize}
        \item Drop $p_{\text{next}}$ on next vertex $v$
        \item $\omega'$ walk of $M_{\text{walk}}$ starting from $v$
        \item Transfer encoding pebbles onto $\omega'$
        \item Placing pebbles at same index
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Recall: Mechanisms of simulation}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices \uncover<2->{\textcolor{green}{{\checkmark}}}
    \item Identify \emph{distinct} vertices of the walk
      \uncover<3->{\textcolor{green}{{\checkmark}}}
    \item Read from and write to simulated tape \uncover<4->{\textcolor{green}{{\checkmark}}}
    \item Preserve tape content through steps of simulated agent
      \uncover<5->{\textcolor{green}{{\checkmark}}}
  \end{itemize}
  \uncover<6->{
    \begin{theorem}
      There is a constant $c\in\mathbb{N}$, such that for any Graph $G$
      with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$,
      there exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates
      the walk of $\mathcal{M}$ or explores $G$.
    \end{theorem}
  }
\end{frame}

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{theorem}
    There is a constant $c\in\mathbb{N}$, such that for any Graph $G$
    with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$,
    there exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates
    the walk of $\mathcal{M}$ or explores $G$.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[below right=0.8 and 1 of d1, rectangle,draw=black] {Apply simulation $\log\log n$ times};
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    Any bounded-degree graph on at most $n$ vertices can be explored using
    $\mathcal{O}(\log\log n)$ pebbles and memory.
  \end{theorem}
\end{frame}

\section{Limitations of pebble machines}
\begin{frame}
  \frametitle{Limitations of pebbles}
  \begin{itemize}
    \item Construct 3-regular graphs agent $\mathcal{A}$ with $p$ pebbles
      cannot traverse
    \item Called $p$-barriers
      \begin{center}
        \resizebox{0.6\textwidth}{!}{\input{tikz/barrierint}}
      \end{center}
    \item $\mathcal{A}$ cannot traverse a $p$-barrier from $v_{1}$ or $v_{2}$
      to $v_{3}$ or $v_{4}$ or vice versa using at most $p$ pebbles
    \item Inductive construction of barriers
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Barrier construction: Initial thoughts}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Subset of pebbles of cardinality $r$
      \item Locality gadget as edges
      \item Pebbles are local: configurations in macro vertex
        $C_{1},\dots,C_{\alpha}$
      \item Project $A$ to agent without pebbles $B$
      \item $B$ has $\alpha$ states and traverses macro vertices
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/projection}}
  \end{columns}
\end{frame}

>>>>

\begin{frame}
  \frametitle{Exploration sequence}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Exploration sequence $e_{1},\dots, e_{n}$
      \item Relative movements of agent
      \item Leaving port: $\ell_{i} + e_{i}\mod d_{v}$
      \item \emph{Universal} for a class of graphs if it explores any graph of
        that class
      \item<2->[$\rightarrow$] Example:
        $\alt<3-3>{\colorbox{green}{1}}{\colorbox{white!0}{1}},
        \alt<4-4>{\colorbox{green}{3}}{\colorbox{white!0}{3}},
        \alt<5-5>{\colorbox{green}{2}}{\colorbox{white!0}{2}},
        \alt<6-6>{\colorbox{green}{2}}{\colorbox{white!0}{2}}$
    \end{itemize}
    \column{0.5\textwidth}
    \begin{center}
      \resizebox{\textwidth}{!}{\input{tikz/moveex}}
    \end{center}
  \end{columns}
\end{frame}

\section{Graph traversal}
\begin{frame}
  \frametitle{Exploring pebble machine}
  \begin{theorem}[Reingold]
    There is an $\mathcal{O}(\log n)$-space algorithm producing a universal
    exploration sequence for any regular graph on $n$ vertices.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \uncover<2->{
        \node[rectangle, draw=green, below right=0.3 and 1 of d1]
        {3-Regularity};
      }
      \uncover<3->{
        \node[rectangle, draw=blue, below right=1 and 1.2 of d1]
        {Periodic Universal Lemma};
      }
      \uncover<4->{
        \node[rectangle, draw=yellow, below right=1.7 and 1 of d1]
        {Closed walk Lemma};
      }
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    There exists a $(\mathcal{O}(1), 0, \mathcal{O}(\log n))$-pebble machine
    that moves along a closed walk and either explores the graph or visits at
    least $n$ distinct vertices, for any graph with bounded degree.
  \end{theorem}
\end{frame}

\begin{frame}
  \frametitle{Closed walk}
  \begin{mdframed}[linecolor=yellow]
    \begin{lemma}[Closed walk]
      An agent following an exploration sequence of the form
      $(e_{1},\dots,e_{k})^{\ast}$ moves along a closed walk.
    \end{lemma}
  \end{mdframed}
  \uncover<2->{
    \begin{mdframed}[linecolor=blue]
      \begin{lemma}[Periodic Universal Lemma]
        There exists an $\mathcal{O}(\log n)$-space algorithm producing a
        universal exploration sequence $(e_{1},\dots,e_{k})^{\ast}$ for any
        $3$-regular graph with at most $n$ vertices.
      \end{lemma}
    \end{mdframed}
  }
\end{frame}

@@

\section{Pebble simulation}
\begin{frame}
  \begin{theorem}
    There is a constant $c\in\mathbb{N}$, such that for any graph $G$
    with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$, there
    exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates the walk
    of $\mathcal{M}$ or explores $G$.
  \end{theorem}
\end{frame}

@@

\begin{frame}
  \frametitle{4. Transfer tape-content through movements}
  \begin{itemize}
    \item Simulated agent moves
    \item Preserve tape-content:
      \begin{itemize}
        \item Drop $p_{\text{next}}$ on next vertex $v$
        \item $\omega'$ walk of $M_{\text{walk}}$ starting from $v$
        \item Transfer encoding pebbles onto $\omega'$
        \item Place pebbles at same index
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Recall: Mechanisms of simulation}
  \begin{itemize}
    \item Simulation of $M_{\text{Walk}}$ which explores at least $2^{m_{1}}$
      distinct vertices \uncover<2->{\textcolor{green}{{\checkmark}}}
    \item Identify \emph{distinct} vertices of the walk
      \uncover<3->{\textcolor{green}{{\checkmark}}}
    \item Read from and write to simulated tape \uncover<4->{\textcolor{green}{{\checkmark}}}
    \item Preserve tape content through steps of simulated agent
      \uncover<5->{\textcolor{green}{{\checkmark}}}
  \end{itemize}
  \uncover<6->{
    \begin{theorem}
      There is a constant $c\in\mathbb{N}$, such that for any graph $G$
      with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$,
      there exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates
      the walk of $\mathcal{M}$ or explores $G$.
    \end{theorem}
  }
\end{frame}

\begin{frame}
  \frametitle{Exploring graphs with $\log\log n$ pebbles}
  \begin{theorem}
    There is a constant $c\in\mathbb{N}$, such that for any graph $G$
    with bounded degree and any $(s,p,2m)$ pebble machine $\mathcal{M}$,
    there exists a $(cs,p+c,m)$ pebble machine $\mathcal{M}'$ that simulates
    the walk of $\mathcal{M}$ or explores $G$.
  \end{theorem}
  \begin{center}
    \begin{tikzpicture}
      \node (d1) {};
      \node[below=2 of d1] (d2) {};
      \draw[->, bend right] (d1) to (d2);
      \node[below right=0.8 and 1 of d1, rectangle,draw=black] {Apply simulation Theorem $\log\log n$ times};
    \end{tikzpicture}
  \end{center}
  \begin{theorem}
    Any bounded-degree graph on at most $n$ vertices can be explored using
    $\mathcal{O}(\log\log n)$ pebbles and memory.
  \end{theorem}
\end{frame}

\section{Limitations of pebble machines}
\begin{frame}
  \frametitle{Limitations of pebbles}
  \begin{itemize}
    \item Construct 3-regular graphs which agent $\mathcal{A}$ with $p$ pebbles
      cannot traverse
    \item Called $p$-barriers
      \begin{center}
        \resizebox{0.6\textwidth}{!}{\input{tikz/barrierint}}
      \end{center}
    \item $\mathcal{A}$ cannot traverse a $p$-barrier from $v_{1}$ or $v_{2}$
      to $v_{3}$ or $v_{4}$ or vice versa using at most $p$ pebbles
    \item Inductive construction of barriers
  \end{itemize}
\end{frame}

@@

\begin{frame}
  \frametitle{Barrier construction: Idea}
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Subset of pebbles of cardinality $r$
      \item Locality gadget as edges
      \item Pebbles are local: configurations in macro vertex
        $C_{1},\dots,C_{\alpha}$
      \item Project $A$ to agent without pebbles $B$
      \item $B$ has $\alpha$ states and traverses macro vertices
    \end{itemize}
    \column{0.5\textwidth}
    \resizebox{\textwidth}{!}{\input{tikz/projection}}
  \end{columns}
\end{frame}
